Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/hat-analysis.tex

HAT systems provide transactions with transactional availability and
either high availability ($1$-available) or sticky high availability
($1$-sticky-available). They offer substantial latency and
availability benefits compared to traditional distributed databases,
yet they cannot achieve all the traditional semantics. In this
section, we delineate ACID, distributed consistency, and session
consistency levels which can be achieved with high availability
(transactional atomicity, variants of Repeatable Read isolation, and
many session guarantees), those with sticky high availability (read
your writes, PRAM and causal consistency) and the properties which
cannot be provided in a HAT system (those preventing Lost Update and
Write Skew, or with recency).  We present a full summary of these
results in Section~\ref{sec:hat-summary}.

@@

Read your writes is not achievable in a highly available
($1$-available) system. Consider a client that executes the following
two transactions in succession:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: r_x(a)
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client executes $T_2$ against the same (partitioned)
server, then it will be able to read its writes. However, if the
network topology shifts and the client can only contact a different
server that is partitioned from the server that executed $T_1$, then
the client will be unable to read its own writes and the system will
have to either stall indefinitely to allow the client to read her
writes (violating transactional availability) or will have to
sacrifice read your writes guarantees. However, if the client remains
sticky with the server that executed $T_1$, then we can disallow this
scenario. Accordingly, read your writes, and, by proxy, causal
consistency and PRAM require stickiness. Read your writes is ``free''
in a sticky system, while the remaining causality and PRAM guarantees
can be accomplished with the algorithms for achieving the remaining
session guarantees.

@@

\vspace{.5em}\noindent{\textbf{Durability}} As we briefly discussed in
Section~\ref{sec:availability}, a client requiring that its
transactions' effects survive $F$ server faults requires at least
($F+1$)-availability.

>>>>

HAT systems provide transactions with transactional availability and
either high availability or sticky high availability. They offer
substantial latency and availability benefits compared to traditional
distributed databases, yet they cannot achieve all the traditional
semantics. In this section, we delineate ACID, distributed
consistency, and session consistency levels which can be achieved with
high availability (transactional atomicity, variants of Repeatable
Read isolation, and many session guarantees), those with sticky high
availability (read your writes, PRAM and causal consistency) and the
properties which cannot be provided in a HAT system (those preventing
Lost Update and Write Skew, or with recency).  We present a full
summary of these results in Section~\ref{sec:hat-summary}.

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions
in succession:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: r_x(a)
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client executes $T_2$ against the same (partitioned)
server, then it will be able to read its writes. However, if the
network topology shifts and the client can only contact a different
server that is partitioned from the server that executed $T_1$, then
the client will be unable to read its own writes and the system will
have to either stall indefinitely to allow the client to read her
writes (violating transactional availability) or will have to
sacrifice read your writes guarantees. However, if the client remains
sticky with the server that executed $T_1$, then we can disallow this
scenario. Accordingly, read your writes, and, by proxy, causal
consistency and PRAM require stickiness. Read your writes is ``free''
in a sticky system, while the remaining causality and PRAM guarantees
can be accomplished with the algorithms for achieving the remaining
session guarantees.

@@

\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
transactions' effects survive $F$ server faults requires at least
$F+1$ correct nodes.
